---
title: "BEE 550 Intro to R"
author: "Jennifer Bradley"
date: "09-24-2024"
output: html_document # Change this to word_document or pdf_document when ready to submit
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)

# don't forget to copy/paste here all of the libraries required for this week's lab
library(tidyverse)
library("ggplot2")
library("gridExtra")

theme_set(theme_bw())
```

## Q1
47

```{r}
12 * 3 - 10 / 2 + 16
```
## Q2
-8

```{r}
12 * (3 - 10 / 2) + 16
```
## Q3
4

```{r}
2 ^ 3 - sqrt(16) + log(1)
```
## Q4
47

```{r}
x = 12 
y = 3
z = 10
w = 2
t = 16

x * y - z / w + t
```
## Q5
The result of exp(log(x)) is always x.

```{r}
x = 6

exp(log(x))
```

## Q6
The area is 6.

```{r}
tri_area = function(L1, L2, L3){
  p = (L1 + L2 + L3) / 2
  area = sqrt(p * (p - L1) * (p - L2) * (p - L3))
  return(area)
}

tri_area(3, 4, 5)
```

## Q7
-2

```{r}
v = c(3, 5, -2, 0, 1)

v[c(3)]
```

## Q8
v[2:4] lists the 2nd, 3rd, and 4th elements of vector v. It seems that the : operator represents "until" or "through." In v[x:y] all elements in the vector v starting with the x'th element until (and including) the y'th element are pulled.

```{r, error = TRUE}
v[2:4]
```

## Q9
The output of mean(v1 + v2) is -1.2
The output of sum(v1 + v2) / length(v1 + v2) is the same, -1.2

```{r}
v1 = v
v2 = c(-5, -1, -9, 2, 0)
print(mean(v1 + v2))

print(sum(v1 + v2) / length(v1 + v2))
```

## Q10

```{r}
my_vector = mean(v1 + v2)

sort(my_vector)
```

## Q11

```{r}
q11_matrix = matrix(c(4, -2, 0, 1, 1, 7), nrow = 3, ncol = 2)
q11_matrix
```

## Q12
The str() function prints the data type of the object fed to the function, and displays the object itself.
```{r}
my_dogs_age = 3.5
my_dogs_name = 'Hobbes'
my_cats_name = NA
is_my_dog_a_puppy = my_dogs_age < 1

mda = str(my_dogs_age)
mdn = str(my_dogs_name)
mcn = str(my_cats_name)
imdap = str(is_my_dog_a_puppy)

```


## Q13
The command is attempting to multiply a non-numeric data type (string) to a numeric type (integer). As this is not possible, R gives the error.

```{r error = TRUE}
my_pet_info =
  list(
    my_dogs_age,
    my_dogs_name,
    my_cats_name,
    is_my_dog_a_puppy
  )

my_dog_info =
  list(
    age = my_dogs_age,
    name = my_dogs_name,
    puppy = is_my_dog_a_puppy
  )

2 * my_dog_info$name

```

## Q14
20.09062
This is the average of all mpg values in the data frame; the average mpg of the cars.

```{r}
mean(mtcars$mpg)
```


## Q15
```{r}
var(mtcars$cyl)
```

## Q16
```{r echo = FALSE}
characters = 
  tibble(
    name = c('Calvin', 'Hobbes', 'Garfield', 'Popeye', 'Bugs Bunny', 'Lisa Simpson', 'Woodstock'),
    species = c('human', 'tiger', 'cat', 'human', 'rabbit', 'human', 'canary'),
    comic_strip = c(TRUE, TRUE, TRUE, TRUE, FALSE, FALSE, TRUE),
    year_created = c(1985, 1985, 1978, 1929, 1940, 1987, 1967)
  )
```

```{r}
characters |>
  arrange(characters, year_created)
```


## Q17

```{r}
my_tibble =
  mtcars |>
  rownames_to_column(var = 'model') |>
  as_tibble() |>
  select(model, mpg, cyl, hp) |>
  mutate(kml = mpg * 0.43) |>
  filter(mpg > 22)
my_tibble
```

## Q18
New_abundance is 16098, the uncertainty is 834

```{r echo = FALSE}
github_address = 'https://github.com/rafaeldandrea/BIO-356/blob/master/gray_whales.xlsx?raw=true'

whale_dataset = 
  rio::import(github_address, range = 'B6:G28') |>
  as_tibble()
```

## Q19
```{r}
plot_whales =
  whale_dataset |>
  ggplot(aes(End_dates, New_abundance)) +
  geom_point() +
  geom_line()
plot_whales
```

## Q20
```{r}
plot_whales =
  whale_dataset |>
  ggplot(aes(End_dates, New_abundance)) +
  geom_line(color = 'green') +
  geom_errorbar(
    aes(
      x = End_dates,
      ymin = New_abundance + New_SE,
      ymax = New_abundance - New_SE
    ),
    color = 'blue'
  ) +
  geom_point(color = 'red') +
  labs(
    x = 'Calendar Year',
    y = 'Population Size'
  ) +
  theme(aspect.ratio = 1) +
  ggtitle('Whale population estimates')
plot_whales
```

## Q21
"Error: unexpected '=' in:
"  print('Two is greater than three')
} else if(2 =" "

```{r error = TRUE}
if(2 > 3){
  print('Two is greater than three')
} else if(2 = 3){ 
  print('Two is equal to three')
} else {
  print('Two is less than three')
}
```

## Q22
```{r}
for(i in 1:5){
  if(i%%2 != 0){
    print(i)
  }
}
```

## Q23
The growth rate is less than 1, so the population declines.

```{r}
lambda = 0.9
N = 50
N_vector = N
for(year in 1:20){
  N = round(lambda * N)
  N_vector = c(N_vector, N)
}

data = 
  tibble(
    year = 0:20,
    abundance = N_vector
  )

plot = 
  data |>
  ggplot(aes(year, abundance)) +
  geom_line(color = 'grey') +
  geom_point()

plot
```

## Q24
The range of the plot with a lower standard deviation is smaller, and the data is more centered around the mean.
```{r}
set.seed(1)

random_numbers1 = rnorm(n = 1000, mean = 0, sd = 1)

random_data1 = 
  tibble(
    x = random_numbers1
  )

plot_data1 = 
  random_data1 |>
  ggplot(aes(x)) +
  geom_histogram(bins = 30, fill = 'darkred') +
  xlim(c(-5, 5))

random_numbers2 = rnorm(n = 1000, mean = 0, sd = 0.4)

random_data2 = 
  tibble(
    x = random_numbers2
  )

plot_data2 =
  random_data2 |>
  ggplot(aes(x)) +
  geom_histogram(bins = 30, fill = 'darkblue') +
  xlim(c(-5, 5))

grid.arrange(plot_data1, plot_data2)
```




