---
title: "BEE 550 Intro to R"
author: "Jennifer Bradley"
date: "09-24-2024"
output: html_document # Change this to word_document or pdf_document when ready to submit
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)

# don't forget to copy/paste here all of the libraries required for this week's lab
library(tidyverse)

theme_set(theme_bw())
```

## Q1
47

```{r}
12 * 3 - 10 / 2 + 16
```
## Q2
-8

```{r}
12 * (3 - 10 / 2) + 16
```
## Q3
4

```{r}
2 ^ 3 - sqrt(16) + log(1)
```
## Q4
47

```{r}
x = 12 
y = 3
z = 10
w = 2
t = 16

x * y - z / w + t
```
## Q5
The result of exp(log(x)) is always x.

```{r}
x = 6

exp(log(x))
```

## Q6
The area is 6.

```{r}
tri_area = function(L1, L2, L3){
  p = (L1 + L2 + L3) / 2
  area = sqrt(p * (p - L1) * (p - L2) * (p - L3))
  return(area)
}

tri_area(3, 4, 5)
```

## Q7
-2

```{r}
v = c(3, 5, -2, 0, 1)

v[c(3)]
```

## Q8
v[2:4] lists the 2nd, 3rd, and 4th elements of vector v. It seems that the : operator represents "until" or "through." In v[x:y] all elements in the vector v starting with the x'th element until (and including) the y'th element are pulled.

```{r, error = TRUE}
v[2:4]
```

## Q9
The output of mean(v1 + v2) is -1.2
The output of sum(v1 + v2) / length(v1 + v2) is the same, -1.2

```{r}
v1 = v
v2 = c(-5, -1, -9, 2, 0)
print(mean(v1 + v2))

print(sum(v1 + v2) / length(v1 + v2))
```

## Q10

```{r}
my_vector = mean(v1 + v2)

sort(my_vector)
```

## Q11

```{r}
q11_matrix = matrix(c(4, -2, 0, 1, 1, 7), nrow = 3, ncol = 2)
q11_matrix
```

## Q12
The str() function prints the data type of the object fed to the function, and displays the object itself.
```{r}
my_dogs_age = 3.5
my_dogs_name = 'Hobbes'
my_cats_name = NA
is_my_dog_a_puppy = my_dogs_age < 1

mda = str(my_dogs_age)
mdn = str(my_dogs_name)
mcn = str(my_cats_name)
imdap = str(is_my_dog_a_puppy)

```


## Q13
The command is attempting to multiply a non-numeric data type (string) to a numeric type (integer). As this is not possible, R gives the error.

```{r error = TRUE}
my_pet_info =
  list(
    my_dogs_age,
    my_dogs_name,
    my_cats_name,
    is_my_dog_a_puppy
  )

my_dog_info =
  list(
    age = my_dogs_age,
    name = my_dogs_name,
    puppy = is_my_dog_a_puppy
  )

2 * my_dog_info$name

```

## Q14
20.09062
This is the average of all mpg values in the data frame; the average mpg of the cars.

```{r}
mean(mtcars$mpg)
```


## Q15
```{r}
var(mtcars$cyl)
```

## Q16

```{r}

```





















